On Sample Size Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-08-29 Manuscript/manuscript.tex
BCA Error

Block bootstrap is a tool to construct confidence intervals (CI) to make
inferences about dependent data. Essentially, it depends on correct estimation
of the uncertainty in the estimation, similar to the standard bootstrap 
\citep{efron1979bootstrap}, but for serially dependent data.
Early ideas of block bootstrap were developed not long after the standard 
bootstrap.\citep{hall1985resampling, carlstein1986use,kunsch1989jackknife} It 
has since
been applied in various fields, for instance, econometrics and 
meteorology.\citep{mackinnon2006bootstrap, varga2017generalised}
Block bootstrap is especially useful for serially dependent data when the serial 
dependence is not specified or not of primary interest. The method is expected 
to produce CIs with coverage rates matching their nominal levels as the sample 
size grows. However, when dealing with finite sample sizes, an important 
question is how large the sample size must be for block bootstrap CIs to have 
the desired coverage rates.
\eds{I think the superscript citations need to go before the punctuation.  
Please check in other AJUR papers and adjust throughout the paper if needed.}

@@

\paragraph{Standard Normal CI}
Assuming that $\hat\theta_n$ is asymptotically normally distributed with 
$\theta$ as the mean, we just need an estimate of the standard error to 
construct an approximate CI.\citep[p.168]{efron1993introduction} Let
$\widehat{\text{SE}}$ be the empirical standard error of the bootstrap point 
estimates $\hat\theta_n^{(b)}$ for $b \in \{1, \ldots, B\}$. Let $z_{(\alpha)}$ 
be the upper $\alpha$ quantile of the standard normal distribution. 
A $(1 - \alpha)100\%$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - z_{(\alpha/2)}\widehat{\text{SE}}).
\]
\eds{If $z_{(\alpha)}$ is the upper quantile, then $z_{(\alpha/2)}$ will be 
positive and $z_{(1-\alpha/2)}$ will be negative for small $\alpha$ values; with
the subtraction of the quantiles, the smaller CI endpoint would appear to
be listed first.  Please check this notation, and the notation for Student's $t$ 
CI, and other intervals which use the $z_{(\alpha)}$ upper standard normal 
quantile.} 
This CI is centered by the original point estimate $\hat\theta_n$ and is 
symmetric. The standard CI is a table-based bootstrap interval approach. Its 
validity relies on whether the distribution of $\hat\theta_n$ is reasonably well 
approximated by its asymptotic normal distribution and whether the bootstrap
$\widehat{\text{SE}}$ approximates the true standard error.

\paragraph{Student's $t$ CI}
The procedure for constructing a Student's $t$ CI based on standard bootstrap is 
described in \citet[p.158]{efron1993introduction}. Let $t_{(\alpha, k)}$ be the 
upper $\alpha$ quantile of a $t$ distribution with $k$ degrees of freedom. With 
block bootstrapping, 
a $(1 - \alpha)100\%$ Student's $t$ CI is
\[
(\hat{\theta}_{n} - t_{(1-\alpha/2), k - 1}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - t_{(\alpha/2), k -1}\widehat{\text{SE}}),
\]
%\eds{define $t_{(.)}$ and check the signs}  \mc{addressed}
where $k$ is the number of blocks. This CI is centered by the original point 
estimate $\hat\theta_n$ and is symmetric. Like the standard normal interval, the
Student's $t$ CI is a table-based bootstrap interval approach. In this case, its 
validity relies on whether the distribution of $\hat\theta_n$ is reasonably well 
approximated by the $t_{k-1}$ distribution with an expected value of $\theta$ 
and whether the bootstrap $\widehat{\text{SE}}$ approximates the true standard 
error.

@@

\jy{I'd slightly increase the heights of the figures such that two of them
  would occupied a whole page. This reduces the complexity of floating in
  finding best locations.}

\jy{Clarify normal margins in captions.}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_mu}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of an AR(1) process with AR coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_sigma}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of an AR(1) process with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. The 
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_phi}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\phi$ of an AR(1) process with
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:phi}
\end{figure}

@@

The simulation design was the same, except the target parameters of $W_t$ were:
1) $\mu = 1$, the mean of $W_t$;
2) $\sigma_w = 1$, the standard deviation of $W_t$; and
3) $\rho$, the lag-1 autocorrelation coefficient of $W_t$.
Although we may have set $\phi$ to these five levels for $X_t$: 
$\{-0.4, -0.2, 0.0, 0.2, 0.4\}$, because of the transformation,
the new five true autocorrelation coefficients for $W_t$ were different.
\mc{I may need some help explaining the calibration of copula parameters}
\jy{see demo code in \texttt{copula.R}}

To calibrate the copula parameters, we first constructed a multivariate
exponential distribution from a normal copula using the \textsl{copula} package.
\citep{nelsen2006introduction} Then we computed $E[XY]$, which 
requires double integration using the \textsl{pracma} package. \citep{pracma} 
Finally, we obtained true $\rho = E[XY] - 1$, which was used as the target
parameter when evaluating CI performance when estimating $\rho$.
After this step, we again plotted empirical coverage 
rates for $\mu$, $\sigma_w$, and $\rho$, as well as 95\% confidence intervals of 
the real coverage. These are displayed in 
\textbf{Figures~\ref{fig:exp_mu}--\ref{fig:exp_rho}}.

\jy{cite the copula book instead of the package} \mc{i cited the nelsen book 
referenced in the manual}
\jy{Cite my book on the copula package}

\jy{Let's call this process marginally transformed AR(1) proess. Itself is not
  an AR(1) process. Correct the captions.}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_mu}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of an AR(1) process with AR coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ transformed to
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_sigma}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of an AR(1) process with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ transformed to
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_rho}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\rho$ of an AR(1) process with
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ transformed to
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_rho}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/exp_rho_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for 
    samples of $n \in \{800, 1600\}$ for an AR(1) process with $\phi = 0.4$ 
    transformed to an exponential distribution iwth $\lambda = 1$. For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:eri}
\end{figure}

It seems that range of sample sizes needed 
to cover parameters of an exponentially distributed AR(1) process is generally 
higher than that needed to cover parameters of a normally
distributed AR(1) process. Despite this, the trends and patterns discussed 
previously with respect
to the performance of different methods and with different parameters are the 
same.
For instance, Student's $t$ CIs still appear to yield higher coverage rates than 
other 
methods. Performance is still generally better when temporal dependence is 
negative rather than positive. Most importantly, percentile, BC, and BCA CIs
still deteriorate in coverage of $\rho$ as sample size increases. As exhibited 
in Figure~\ref{fig:eri}, percentile and BC CIs still have the same bias
issue. Overall, the results with an underlying marginal exponential distribution 
are very similar to the results with an underlying normal distribution from the
main study.

>>>>

Block bootstrap is a tool to construct confidence intervals (CI) to make
inferences about dependent data. Essentially, it depends on correct estimation
of the uncertainty in the estimation, similar to the standard bootstrap 
\citep{efron1979bootstrap}, but for serially dependent data.
Early ideas of block bootstrap were developed not long after the standard 
bootstrap.\citep{hall1985resampling, carlstein1986use,kunsch1989jackknife} It 
has since
been applied in various fields, for instance, econometrics and 
meteorology.\citep{mackinnon2006bootstrap, varga2017generalised}
Block bootstrap is especially useful for serially dependent data when the serial 
dependence is not specified or not of primary interest. The method is expected 
to produce CIs with coverage rates matching their nominal levels as the sample 
size grows. However, when dealing with finite sample sizes, an important 
question is how large the sample size must be for block bootstrap CIs to have 
the desired coverage rates.
\eds{I think the superscript citations need to go before the punctuation.  
Please check in other AJUR papers and adjust throughout the paper if needed.}
\mc{from the style guide: "Please bold and superscript the numbers for 
references in the text.1 Separate superscripted numbers by comma and space,1, 2 
they should be separated by an en-dash if the consecutive list of more than two 
numbers is used.1–3 List them AFTER punctuation (be it comma or period) with no 
space.2}

@@

\paragraph{Standard Normal CI}
Assuming that $\hat\theta_n$ is asymptotically normally distributed with 
$\theta$ as the mean, we just need an estimate of the standard error to 
construct an approximate CI.\citep[p.168]{efron1993introduction} Let
$\widehat{\text{SE}}$ be the empirical standard error of the bootstrap point 
estimates $\hat\theta_n^{(b)}$ for $b \in \{1, \ldots, B\}$. Let $z_{(\alpha)}$ 
be the quantile function $F^{-1}(\alpha)$ of the standard normal distribution. 
A $(1 - \alpha)100\%$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - z_{(\alpha/2)}\widehat{\text{SE}}).
\]
\eds{If $z_{(\alpha)}$ is the upper quantile, then $z_{(\alpha/2)}$ will be 
positive and $z_{(1-\alpha/2)}$ will be negative for small $\alpha$ values; with
the subtraction of the quantiles, the smaller CI endpoint would appear to
be listed first.  Please check this notation, and the notation for Student's $t$ 
CI, and other intervals which use the $z_{(\alpha)}$ upper standard normal 
quantile.} 
\mc{Replaced upper quantile with quantile function}
This CI is centered by the original point estimate $\hat\theta_n$ and is 
symmetric. The standard CI is a table-based bootstrap interval approach. Its 
validity relies on whether the distribution of $\hat\theta_n$ is reasonably well 
approximated by its asymptotic normal distribution and whether the bootstrap
$\widehat{\text{SE}}$ approximates the true standard error.

\paragraph{Student's $t$ CI}
The procedure for constructing a Student's $t$ CI based on standard bootstrap is 
described in \citet[p.158]{efron1993introduction}. Let $t_{(\alpha, k)}$ be the 
quantile function $F^{-1}(\alpha, k)$ of a $t$ distribution with $k$ degrees of 
freedom. With 
block bootstrapping, 
a $(1 - \alpha)100\%$ Student's $t$ CI is
\[
(\hat{\theta}_{n} - t_{(1-\alpha/2), k - 1}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - t_{(\alpha/2), k -1}\widehat{\text{SE}}),
\]
%\eds{define $t_{(.)}$ and check the signs}  \mc{addressed}
where $k$ is the number of blocks. This CI is centered by the original point 
estimate $\hat\theta_n$ and is symmetric. Like the standard normal interval, the
Student's $t$ CI is a table-based bootstrap interval approach. In this case, its 
validity relies on whether the distribution of $\hat\theta_n$ is reasonably well 
approximated by the $t_{k-1}$ distribution with an expected value of $\theta$ 
and whether the bootstrap $\widehat{\text{SE}}$ approximates the true standard 
error.

@@

\jy{I'd slightly increase the heights of the figures such that two of them
  would occupied a whole page. This reduces the complexity of floating in
  finding best locations.}
\mc{addressed}

\jy{Clarify normal margins in captions.}
\mc{addressed}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_mu}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of an AR(1) process with an underlying standard 
    normal distribution with AR coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_sigma}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of an AR(1) process with an 
    underlying standard normal distribution with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. The 
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_phi}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\phi$ of an AR(1) process with 
    an underlying standard normal distribution with 
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:phi}
\end{figure}

@@

The simulation design was the same, except the target parameters of $W_t$ were:
1) $\mu = 1$, the mean of $W_t$;
2) $\sigma_w = 1$, the standard deviation of $W_t$; and
3) $\phi$, the lag-1 autocorrelation coefficient of $W_t$.
Let $\rho$ be the lag-1 autocorrelation coefficient of $X_t$.
Although we may have set $\rho$ to these five levels for $X_t$: 
$\{-0.4, -0.2, 0.0, 0.2, 0.4\}$, because of the transformation,
the new five true autocorrelation coefficients for $W_t$ were different.
\mc{I may need some help explaining the calibration of copula parameters}
\jy{see demo code in \texttt{copula.R}}

To calibrate the copula parameters, we first constructed a multivariate
exponential distribution from a normal copula with 
$\rho \in \{-0.4, -0.2, 0.0, 0.2, 0.4\} $using the \textsl{copula} package.
\citep{hofert2018elements} Then we computed $E[XY]$, which 
requires double integration using the \textsl{pracma} package. \citep{pracma} 
Finally, we obtained true $\phi = E[XY] - 1$, which was used as the target
parameter when evaluating CI performance when estimating $\phi$.
After this step, we again plotted empirical coverage 
rates for $\mu$, $\sigma_w$, and $\phi$, as well as 95\% confidence intervals of 
the real coverage. These are displayed in 
\textbf{Figures~\ref{fig:exp_mu}--\ref{fig:exp_phi}}.

\jy{cite the copula book instead of the package} \mc{i cited the nelsen book 
referenced in the manual}
\jy{Cite my book on the copula package}
\mc{addressed}

\jy{Let's call this process marginally transformed AR(1) proess. Itself is not
  an AR(1) process. Correct the captions.}
\mc{addressed}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_mu}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of a marginally transformed AR(1) process with an 
    underlying exponential distribution ($\lambda = 1$) with AR 
    coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ with
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_sigma}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of a marginally transformed AR(1) 
    process with an 
    underlying exponential distribution ($\lambda = 1$) with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ transformed to
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_phi}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\phi$ of a marginally 
    transformed AR(1) process with 
    an 
    underlying exponential distribution ($\lambda = 1$) with
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ transformed to
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_phi}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/exp_pho_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for 
    samples of $n \in \{800, 1600\}$ for an AR(1) process with $\phi = 0.4$ 
    transformed to an exponential distribution iwth $\lambda = 1$. For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:eri}
\end{figure}

It seems that range of sample sizes needed 
to cover parameters of an exponentially distributed AR(1) process is generally 
higher than that needed to cover parameters of a normally
distributed AR(1) process. Despite this, the trends and patterns discussed 
previously with respect
to the performance of different methods and with different parameters are the 
same.
For instance, Student's $t$ CIs still appear to yield higher coverage rates than 
other 
methods. Performance is still generally better when temporal dependence is 
negative rather than positive. Most importantly, percentile, BC, and BCA CIs
still deteriorate in coverage of $\phi$ as sample size increases. As exhibited 
in Figure~\ref{fig:eri}, percentile and BC CIs still have the same bias
issue. Overall, the results with an underlying marginal exponential distribution 
are very similar to the results with an underlying normal distribution from the
main study.
